A Type System for Safe and Scalable Parallelism using Logical Regions

### 2012-07-05 doc/popl2013/evaluation.tex
Messing with figure sizes in the evaluation
Working some more on the evaluation

To determine the benefits of the Legion type system we evaluate
it on three criteria:
\begin{itemize}
\item Expressivity: can real applications be expressed (Section \ref{subsec:expressivity})
\item Overhead: can the it reduce checking overhead (Section \ref{subsec:overhead})
\item Scalability: can it leverage hierarchical scheduling to scale (Section \ref{subsec:scalability})
\end{itemize}
Our evaluation uses a prototype implementation of the Legion langauge and programming model.
The prototype consists of two components: an implementation of a type checker
for the langauge introduced in Section \ref{subsec:langdef} and a C++ runtime library
for executing programs written in the Legion programming model\cite{Legion12}.  All experiments
are conducted on the Keeneland supercomputer\cite{Keeneland}.  Each node of the Keeneland
machine consists of two Xeon 5660 processors, three Tesla M2090 GPUs, and 24 GB of DRAM.  Nodes
are connected by a QDR Infiniband interconnect.

@@

\subsection{Checking Overhead}
\label{subsec:overhead}
Our initial implementation of Legion consisted only of a C++ library for Legion 
programs\cite{Legion12} which contained no checking of region memory accesses.  In the process
of implementing applications in Legion we frequently encountered memory corruption due to
illegal region accesses because of bugs and the lack of checking.  Many times these corruptions 
occurred on remote nodes or on GPUs.  Finding these bugs required either multiple {\tt gdb} windows
for observing execution on multiple nodes simultaneously or using {\tt printf} for debugging on
GPUs.  To alleviate this problem we added a feature to our runtime system that dynamically checked region 
accesses for both CPUs and GPUs to guarantee safety.  However, in many cases this 
added considerable overhead.

To avoid the overhead of dynamic checks for all memory accesses we implemented a type checker of the
language in Section \ref{subsec:langdef}.  Using this checker we were able to implement programs
in this language and type check them.  The soundness proof for the type system allowed us to safely
elide all dynamic checks of region accesses.

@@

\begin{figure}
\begin{center}
\includegraphics[scale=0.40]{figs/circuit_96_popl.pdf}
\end{center}
\caption{Overhead for the Circuit simulation with 96 total pieces.\label{fig:ckt_overhead}}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[scale=0.40]{figs/fluid_19200_popl.pdf}
\end{center}
\caption{Overhead for the Fluid simulation with 19200 cells.\label{fig:fluid_overhead}}
\end{figure}

Figures \ref{fig:ckt_overhead}, \ref{fig:fluid_overhead}, and \ref{fig:amr_overhead} show 
the total time spent by all CPUs and GPUs in each phase of the application.  In 
each figure the problem size stays the same as the number of processors are increased
(strong scaling).  Figure \ref{fig:amr_overhead} includes multiple problem sizes to show
how overhead is affected by changing problem size (weak scaling).  For cases where there
is an existing implementation to compare against we have included a dotted line indicating
baseline performance.  In a few cases (e.g. Figure \ref{fig:amr4096}) checking overhead is the difference
between better and worse performance than the baseline.  For an explanation of overall
performance relative to the baseline implementations please refer to \cite{Legion12}.

\begin{figure}
\begin{center}
\subfigure[4096 cells per dimension]
{
\includegraphics[scale=0.40]{figs/amr_4096_popl.pdf}
\label{fig:amr4096}
}
\subfigure[8192 cells per dimension]
{
\includegraphics[scale=0.40]{figs/amr_8192_popl.pdf}
\label{fig:amr8192}
}
\subfigure[16384 cells per dimension]
{
\includegraphics[scale=0.40]{figs/amr_16384_popl.pdf}
\label{fig:amr16384}
}
\end{center}
\caption{Overhead for the AMR application across problem sizes.\label{fig:amr_overhead}}
\end{figure}

>>>>

To determine the benefits of the Legion type system we evaluate
it on three criteria:
\begin{itemize}
\item Expressivity: can real applications be expressed (Section \ref{subsec:expressivity})
\item Overhead: can it reduce checking overhead (Section \ref{subsec:overhead})
\item Scalability: can it leverage hierarchical scheduling to scale (Section \ref{subsec:scalability})
\end{itemize}
Our evaluation uses a prototype implementation of the Legion langauge and programming model.
The prototype consists of two components: an implementation of a type checker
for the langauge introduced in Section \ref{subsec:langdef} and a C++ runtime library
for executing programs written in the Legion programming model\cite{Legion12}.  All experiments
are conducted on the Keeneland supercomputer\cite{Keeneland}.  Each node of the Keeneland
machine consists of two Xeon 5660 processors, three Tesla M2090 GPUs, and 24 GB of DRAM.  Nodes
are connected by a QDR Infiniband interconnect.

@@

\subsection{Checking Overhead}
\label{subsec:overhead}
Our initial implementation of Legion consisted only of a C++ library for Legion 
programs\cite{Legion12} which contained no checking of region memory accesses.  In the process
of implementing applications in Legion we frequently encountered memory corruption due to
illegal region accesses caused by bugs and a lack of checking.  Many times these corruptions 
occurred on remote nodes or on GPUs.  Finding these bugs required either multiple {\tt gdb} sessions 
for observing execution on multiple nodes simultaneously or using {\tt printf} for debugging on
GPUs.  To alleviate this problem we added a feature to our runtime system that dynamically checked all region 
accesses for both CPUs and GPUs to guarantee safety.  However, this feature
added considerable runtime overhead.

To avoid the overhead of dynamic checks for all memory accesses we implemented a type checker of the
language in Section \ref{subsec:langdef}.  Using the type checker we were able to statically verify
the correctness of all region accesses and elide all the dynamic checks.

@@

\begin{figure}
\begin{center}
\includegraphics[scale=0.35]{figs/circuit_96_popl.pdf}
\end{center}
\vspace{-2mm}
\caption{Overhead for the Circuit simulation with 96 total pieces.\label{fig:ckt_overhead}}
\vspace{-6mm}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[scale=0.35]{figs/fluid_19200_popl.pdf}
\end{center}
\vspace{-2mm}
\caption{Overhead for the Fluid simulation with 19200 cells.\label{fig:fluid_overhead}}
\vspace{-6mm}
\end{figure}

Figures \ref{fig:ckt_overhead}, \ref{fig:fluid_overhead}, and \ref{fig:amr_overhead} show 
the total time spent by all CPUs and GPUs in each phase of the application.  The topmost
component of each bar shows the overhead of the dynamic checks.  In 
each figure the problem size stays the same as the number of processors are increased
(strong scaling).  Figure \ref{fig:amr_overhead} includes multiple problem sizes to show
how overhead is affected by changing problem size (weak scaling).  For cases where there
is an existing implementation to compare against we have included a dotted line indicating
baseline performance.  In a few cases (e.g. Figure \ref{fig:amr4096}) checking overhead is the difference
between better and worse performance than the baseline.  For an explanation of overall
performance relative to the baseline implementations please refer to \cite{Legion12}.

\begin{figure}
\begin{center}
\subfigure[4096 cells per dimension]
{
\includegraphics[scale=0.35]{figs/amr_4096_popl.pdf}
\label{fig:amr4096}
}
\subfigure[8192 cells per dimension]
{
\includegraphics[scale=0.35]{figs/amr_8192_popl.pdf}
\label{fig:amr8192}
}
\subfigure[16384 cells per dimension]
{
\includegraphics[scale=0.35]{figs/amr_16384_popl.pdf}
\label{fig:amr16384}
}
\end{center}
\vspace{-2mm}
\caption{Overhead for the AMR application for three different problem sizes.\label{fig:amr_overhead}}
\vspace{-6mm}
\end{figure}
